\LARGE \bf Mean Estimation from One-Bit Measurements

### 2021-01-06 adaptive-estimation.tex
minor changes to proofs of Sec IV
fixed 1/2 typo and \pm in the proof of thm:sgd
ready for submission

We begin with our first theorem, whose proof we provide in
Appendix~\ref{proof:thm:adpative_lower_bound}.
\begin{thm}[Fundamental limits]\label{thm:adpative_lower_bound}
  Let Assumption~\ref{assump:failure_rate} hold.
  Let ${\theta}_n$ be any estimator of $\theta$ in the adaptive setting of
  Figure~\ref{fig:setup}(ii). Assume that the prior
  density $\pi(\cdot)$ on $\theta$ converges to zero
  at the endpoints of the interval $\Theta$ and
  define the prior Fisher information
  $I_0 \defeq \E_\pi[(\pi'(\theta) / \pi(\theta))^2]$.
  Then
  \begin{equation*}
    \E\left[ (\theta-{\theta}_n)^2 \right] \geq   \frac{n}{ 4f^2(0) n + I_0}.
  \end{equation*}
\end{thm}

%We first bound from above the Fisher information of any set of $n$ one-bit messages with respect to $\theta$. The result then follows using the van-Trees inequality which bounds from below the risk of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are in

%% Theorem~\ref{thm:adpative_lower_bound} implies that any estimator ${\theta}_n$ from any adaptive encoding scheme satisfies
%% \begin{equation*}
%% n\mathbb E\left[ (\theta-\theta_n)^2 \right] \geq  \frac{1}{4f^2(0)}+O(n^{-1}),
%% \end{equation*}
%% and 
%% \begin{equation*}
%% \ARE({\theta}_n) \leq 4f^2(0)\sigma^2 = \eta(0)\sigma^2.
%% \end{equation*}
It is possible to extend Theorem~\ref{thm:adpative_lower_bound} to other
loss functions using a more subtle version of the Van Trees
inequality~\cite{efroimovich1980information}; see also
\cite{DBLP:journals/corr/abs-1902-08582}.

@@

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(0)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    Let $\{P_\theta\}_{\theta \in \R}$ be the family of distributions
    with density $f(\cdot - \theta)$, where $f$ has median 0.
    Let $h_n \to h \in \R$, and define the distributions
    $P_n = P_{\theta + h_n/\sqrt{n}}^n$. Then
    \begin{equation*}
      \sqrt{n}\left(\bar{\theta}_n - \theta - h_n / \sqrt{n}\right)
      \mathop{\cd}_{P_n}
      \normal\left(0, \frac{1}{4 f(0)^2}\right),
    \end{equation*}
    and for any bounded, symmetric, and quasi-convex function $L$,
    \begin{align} 
      & \sup_{c < \infty} \limsup_{n \to \infty}
      \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}
      \E_\tau \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[L(W/ 2 f(0)) \right],
        \label{eq:attaining_LAM}
    \end{align}
    where $W \sim \normal(0,1)$. 
    %
    %
  \item \label{item:sgd-ms-convergence} Assume the stepsizes $\gamma_n$
    satisfy both conditions~\eqref{eqn:lazy-gamma}
    and~\eqref{eqn:stringent-gamma}. Let
    $\pi$ be a distribution on $\R$. Then
    \begin{equation*}
      \int \E\Big[( \bar{\theta}_n - \theta )^2\Big] \pi(d\theta)
      = \frac{1}{4 n f(0)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}

\noindent
We provide the proofs of the items in Appendices~\ref{sec:proof-normal-sgd},
\ref{sec:proof-sgd-regular}, \ref{sec:proof-sgd-ms-convergence},
respectively.

@@

\end{tikzpicture}
\end{center}
\caption{Distributed encoding with one round of interaction: The estimation obtained from the first $n_1$ bits in a distributed manner is utilized in obtaining another $n-n_1$ bits in a distributed manner. 
\label{fig:one_round}
}
\end{figure}

@@

@@

\addplot[color = red, solid, smooth] plot table [x = itr, y = split, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
\addlegendentry{single interaction};

>>>>

We begin with our first theorem, whose proof we provide in
Appendix~\ref{proof:thm:adpative_lower_bound}.
\begin{thm}[Fundamental limits]\label{thm:adpative_lower_bound}
  Let Assumption~\ref{assump:failure_rate} hold.
  Let ${\theta}_n$ be any estimator of $\theta$ in the adaptive setting of
  Figure~\ref{fig:setup}(ii). Assume that the prior
  density $\pi(\cdot)$ on $\theta$ converges to zero
  at the endpoints of the interval $\Theta$ and
  define the prior Fisher information
  $I_0 \defeq \E_\pi[(\pi'(\theta) / \pi(\theta))^2]$.
  Then
  \begin{equation*}
    \E\left[ (\theta-{\theta}_n)^2 \right] \geq   \frac{1}{ 4f^2(0) n + I_0}.
  \end{equation*}
\end{thm}
%Theorem~\ref{thm:adpative_lower_bound} can be extended to other loss functions using more general versions of the Van Trees inequality~\cite{efroimovich1980information}; see also \cite{DBLP:journals/corr/abs-1902-08582}.

@@

Then we can adapt the results of Polyak and Juditsky~\cite{PolyakJu92}
on the asymptotic normality of averaged stochastic gradient estimators
to establish the following theorem.
\begin{thm}
  \label{thm:sgd}
  Define the average $\bar{\theta}_n \defeq \frac{1}{n}
  \sum_{i = 1}^n \theta_i$. Assume
  that in a neighborhood
  of $\theta = \mbox{med}(P)$,
  the distribution $P$ has a Lipschitz continuous density $f$.
  Then
  \begin{enumerate}[(i)]
  \item \label{item:normal-sgd}
    Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
    Then
    \begin{equation*}
      \sqrt{n}\left( \bar{\theta}_n - \theta\right)
      \cd \normal\left(0,\frac{1}{4 f(0)^2}\right).
    \end{equation*}
    %
    %
  \item \label{item:sgd-regular}
    %(or that $f(x-\theta)$ is differentiable in quadratic mean).  
    Let $\{P_\theta\}_{\theta \in \R}$ be the family of distributions
    with density $f(\cdot - \theta)$, where $f$ has median 0.
    Let $h_n \to h \in \R$, and define the distributions
    $P_n = P_{\theta + h_n/\sqrt{n}}^n$. Then
    \begin{equation*}
      \sqrt{n}\left(\bar{\theta}_n - \theta - h_n / \sqrt{n}\right)
      \mathop{\cd}_{P_n}
      \normal\left(0, \frac{1}{4 f(0)^2}\right),
    \end{equation*}
    and for any bounded, symmetric, and quasi-convex function $L$,
    \begin{align} 
      & \sup_{c < \infty} \limsup_{n \to \infty}
      \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}
      \E_\tau \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
      \\
      & \qquad \qquad \qquad \qquad = \mathbb E \left[L(Z/ 2 f(0)) \right],
        \label{eq:attaining_LAM}
    \end{align}
    where $Z \sim \normal(0,1)$. 
    %
    %
  \item \label{item:sgd-ms-convergence} Assume the stepsizes $\gamma_n$
    satisfy both conditions~\eqref{eqn:lazy-gamma}
    and~\eqref{eqn:stringent-gamma}. Let
    $\pi$ be a distribution on $\R$. Then
    \begin{equation*}
      \int \E\Big[( \bar{\theta}_n - \theta )^2\Big] \pi(d\theta)
      = \frac{1}{4 n f(0)^2} + o(n^{-1}). 
    \end{equation*}
  \end{enumerate}
\end{thm}

\noindent
We provide the proofs of items (i)-(iii) in Appendices~\ref{sec:proof-normal-sgd},
\ref{sec:proof-sgd-regular}, \ref{sec:proof-sgd-ms-convergence},
respectively.

@@

\end{tikzpicture}
\end{center}
\caption{Distributed encoding with one round of threshold adaptation. The estimation obtained from the first $n_1$ bits in a distributed manner is utilized in obtaining another $n-n_1$ bits in a distributed manner. 
\label{fig:one_round}
}
\end{figure}

@@

Figure \ref{fig:adaptive_error} illustrates the empirical risks 
of the estimator \eqref{eq:sgd_alg} and an estimator obtained using one round of threshold adaptation under a series of Monte Carlo simulations when $f(x)$ is the standard normal desnity.

@@

\addplot[color = red, solid, smooth] plot table [x = itr, y = split, col sep=comma] {./SimRes/sim_res_nMonte5000.csv};
\addlegendentry{one adaptation};

### 2021-01-06 distributed-estimation.tex
ready for submission

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
  \begin{center}
\input{Figs/opt_thr_den.tex}
\caption{\label{fig:minimax_support}
Optimal threshold density. The threshold density $\lambda^\star(dt)$ (blue) that maximizes the asymptotic relative efficiency for $f(x) = \Ncal(\theta,\sigma^2)$ and $\theta \in [-T,T]$.
%
The continuous curve (red) is the ARE for each $\theta \in [-1/2,1/2]$ under the optimal density, hence the minimax ARE is the minimal value of this curve. The dashed curve (green) is the ARE when the threshold values are uniformly distributed over $[-1/2,1/2]$; its minimal value is $\kappa_{\unif}$ \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

\addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{uniform threshold desnity};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma$ for $f(x) = \Ncal(\theta,\sigma^2)$ and $\theta \in \Theta = [-1/2,1/2]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$ of \eqref{eq:uniform_risk}.}.
\end{center}
\end{figure}

>>>>

\begin{figure}
  %%%%% COMMENTED TIKZ PICTURE %%%%%
  \begin{center}
\input{Figs/opt_thr_den.tex}
\caption{\label{fig:minimax_support}
Optimal threshold density. The threshold density $\lambda^\star(dt)$ (blue) that maximizes the asymptotic relative efficiency for $f(x)$ the normal desnity with variance $\sigma^2$ and $\Theta= [-1/2,1/2]$. 
%
The continuous curve (red) is the ARE for each $\theta \in [-1/2,1/2]$ under the optimal density, hence the minimax ARE is the minimal value of this curve. The dashed curve (green) is the ARE when the threshold values are uniformly distributed over $[-1/2,1/2]$; its minimal value is $\kappa_{\unif}$ \eqref{eq:uniform_risk}. }
\end{center}
\end{figure}

@@

\addplot[color = black!35!green, smooth, dashed, line width = 1pt] plot table [x = x, y = z, col sep=comma] {./Figs/minmax_ARE_b0.5.csv};
\addlegendentry{uniform threshold desnity};
\end{axis}
\end{tikzpicture}
\caption{\label{fig:minimax_ARE} Minimax ARE versus $\sigma$ for $f(x)$ the standard normal density with variance $\sigma^2$ and $\Theta = [-1/2,1/2]$. The dashed curve (green) is the ARE under a uniform threshold density over $\Theta$ given by $K_{\unif}\sigma^2$ of \eqref{eq:uniform_risk}.}.
\end{center}
\end{figure}

### 2021-01-06 onebit_full.tex
Rewriting intro
Working on problem setting
Compiler
Fixed proof of (previously) Lemma 15 and made changes in structure regarding this lemma. Now one lemma (Lemma 12) is used in three different results.
Corrected proof of isoprimetric lemma
Working on proof of SGD
Working on rewriting proof of theorem 2
Giving the one step estimator proof
Starting to work on distributed part
Rewrote entire proof of local asymptotic normality
Finished writing major proofs and fixing references. Still needs a once through
added funding information
ready for submission

%% \documentclass[11pt]{article}
\documentclass[letterpaper, 11pt]{IEEEtran}      % Use this line for a4 paper

@@

\author{
    \IEEEauthorblockN{
Alon Kipnis\IEEEauthorrefmark{1} and   
    John C. Duchi\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}  \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Stanford University, Department of Statistics \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Stanford University, Department of Electrical Engineering. }
}

@@

%{\color{red}  See LASSO results in http://arxiv.org/abs/1506.02181v1}

@@

%% \input{adaptive-estimation}

%% \input{distributed-estimation}

%% \input{conclusions}

%\section*{ACKNOWLEDGMENT}
%This research is supported in parts by...
%Tsachy Weissman, Stefano Rini, Robert Gray.

@@

\appendix

%% \input{appendices}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtran}
\bibliography{IEEEfull,onebit}

>>>>

%% \documentclass[11pt]{article}
\documentclass[letterpaper, 11pt, draftcls]{IEEEtran}      % Use this line for a4 paper

@@

\author{
    \IEEEauthorblockN{
Alon Kipnis\IEEEauthorrefmark{1} and   
    John C. Duchi\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}  \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Stanford University, Department of Statistics \\}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Stanford University, Department of Electrical Engineering. }

\thanks{This paper was presented in part at the 55th Annual Allerton Conference on Communication, Control, and Computing (Allerton) \cite{KipnisAllerton2017}.
}
}

%% \renewenvironment{figure}{%
%%   \ignorespaces
%% }{}

@@

@@

\input{adaptive-estimation}

\input{distributed-estimation}

\input{conclusions}

\section*{Acknowledgments}
%% \textbf{WRITE THIS SECTION}
The work of A. Kipnis was supported in part by funding from the NSF under Grant No.~DMS-1418362 and DMS-1407813, and by a fellowship from the Koret Foundation.

@@

\newpage
% \appendix

\section*{Appendices}

\input{uniform-are-weirdos}
\input{proof-adaptive-optimality}
\input{proof-sgd}
\input{proof-lan-distributed}
\input{proof-no-distributed}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{IEEEtran}
\bibliography{IEEEfull,onebit,bib}
